---
title: "1. exploratory analyses"
author: "Dan Weinberger"
date: '2023-01-05'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(arrow)
library(tidyr)
library(dplyr)
library(ggplot2)
source('./R/schema.R')

```

## Read in the parquet database

```{r} 
pq_stream_denom <- open_dataset("T:/parquet/denom/", format = "parquet", schema = sch_denom)
```

```{r} 
pq_stream_inp <- open_dataset("T:/parquet/inp/", format = "parquet", schema = sch_inp_basek)
```

## Test databases 

Take a test pull from the dataset
```{r}
test1 <- pq_stream_denom %>% 
  slice_head( n = 10000) %>%
  collect() #use collect() to take parquet to dataframe
```

Look at age distribution of all enrollees in 2015
should be ~54 million enrollees in 2015 (58.6 million in 2021)..

to run on the csv_stream takes 151 seconds
to run on the parquet data stream takes 22 seconds
```{r}
ptm <- proc.time()
N_age2 <- pq_stream_denom %>% 
  filter(AGE>=65 & RFRNC_YR==15) %>%
  select(AGE,RFRNC_YR) %>%
group_by(AGE) %>%
  summarize(N_obs = n()) %>% #use collect() to take parquet to dataframe
    collect()

proc.time() - ptm

ggplot(N_age2, aes(x=AGE, y=N_obs)) +
  geom_line()+
  theme_classic()
```


```{r}
test <-pq_stream_inp %>%
slice_head( n = 50) %>%
  collect() 
```

Number of admissions by week, nationally

Break out by week and ICD chapter AND region
```{r,eval=F}
#runs in 2.5 min
ptm <- proc.time()
N_chapter_date_state <- pq_stream_inp %>%
  mutate(princ_icd_chapt =substr(PRNCPAL_DGNS_CD,1,1)) %>%
  group_by(princ_icd_chapt,ADMSN_DT,STATE_CD) %>%
  summarize(N_cases = n()) %>%
  collect() %>%
  mutate(date=as.Date(ADMSN_DT,'%Y%m%d')) 
saveRDS(N_chapter_date_state,"T:/results/inp/N_chapter_date_state.rds")
proc.time() - ptm
```


```{r}
state_cw <- read.csv('./Data/SSA_state_crosswalk.csv')

readRDS("T:/results/inp/N_chapter_date_state.rds")

state.info <- cbind.data.frame(state.name, state.abb, state.region)

statedate <- N_chapter_date_state %>%
  filter(date>='2017-01-01' &date<='2021-11-20' )%>%
  mutate(STATE_CD=as.numeric(STATE_CD)) %>%
  left_join(state_cw, by=c('STATE_CD'='STATE_SSA')) %>%
left_join(state.info,by=c('STATENAME'='state.name')) %>%
  mutate(weekdate=lubridate::floor_date(date, 'week')) %>%
  group_by(princ_icd_chapt,state.region,weekdate) %>%
  summarize(N_cases=sum(N_cases)) %>%
  ungroup()
```

National, all causes
```{r}
statedate%>%
   group_by(weekdate) %>%
  summarize(N_cases=sum(N_cases)) %>%
    ggplot(aes(x=weekdate, y=N_cases))+
  geom_line()+
  theme_classic()+
  ylim(0,NA) +
  geom_vline(xintercept=as.Date(c('2020-03-15','2020-12-15','2021-08-15')), lty=2, col='gray')
```
National, by chapter
```{r, fig.width=10, fig.height=9}
statedate%>%
   group_by(princ_icd_chapt,weekdate) %>%
  summarize(N_cases=sum(N_cases)) %>%
    ggplot(aes(x=weekdate, y=N_cases))+
  facet_wrap(~princ_icd_chapt, scales='free_y') +
  geom_line()+
  theme_classic()+
  ylim(0,NA) +
  geom_vline(xintercept=as.Date(c('2020-03-15','2020-12-15','2021-08-15')), lty=2, col='gray')
```

Sam but filter by region (Northeast, North Central, South, West
)
```{r}
statedate%>%
  filter(state.region=='West') %>%
  ggplot(aes(x=weekdate, y=N_cases))+
  geom_line()+
  facet_wrap(~princ_icd_chapt, scales='free_y') +
  theme_classic()+
  ylim(0,NA) +
  geom_vline(xintercept=as.Date(c('2020-03-15','2020-12-15','2021-08-15')), lty=2, col='gray')

```



```{r}

ptm <- proc.time()
inp1 <- pq_stream_inp %>%
  mutate(date=as.Date(ADMSN_DT,'%Y%m%d'),     
             all_dx = paste0(ICD_DGNS_CD1,
                         ICD_DGNS_CD1,
                         ICD_DGNS_CD2,
                         ICD_DGNS_CD3,
                         ICD_DGNS_CD4,
                         ICD_DGNS_CD5,
                         ICD_DGNS_CD6,
                         ICD_DGNS_CD7,
                         ICD_DGNS_CD8,
                         ICD_DGNS_CD9,
                         ICD_DGNS_CD10,
                         ICD_DGNS_CD11,
                         ICD_DGNS_CD12,
                         ICD_DGNS_CD13,
                         ICD_DGNS_CD14,
                         ICD_DGNS_CD15,
                         ICD_DGNS_CD16,
                         ICD_DGNS_CD17,
                    ICD_DGNS_CD18,
                    ICD_DGNS_CD19,
                    ICD_DGNS_CD20,
                    ICD_DGNS_CD21,
                    ICD_DGNS_CD22,
                    ICD_DGNS_CD23,
                    ICD_DGNS_CD24,
                    ICD_DGNS_CD24)
                  )%>%
  filter( grepl('J12',all_dx)|grepl('J13',all_dx)|grepl('J14',all_dx)|grepl('J15',all_dx)|grepl('J16',all_dx)|grepl('J127',all_dx)|grepl('J18',all_dx)) %>%
  group_by(date) %>%
  summarize(J12_18=n()) %>%
  collect()
proc.time() - ptm

inp1 %>%
  arrange(date) %>%
  filter(date>='2017-01-01') %>%
  ggplot(aes( x=date, y=J12_18) )+
  geom_line()+
  theme_classic()
  
```


```{r}
ptm <- proc.time()
N_age_die <- pq_stream_denom %>%
  filter(DEATH_DT != '') %>%
  select(AGE,RFRNC_YR) %>%
  filter(RFRNC_YR==15) %>%
group_by(AGE) %>%
  summarize(N_deaths = n()) %>% #use collect() to take parquet to dataframe
    collect()

proc.time() - ptm

ggplot(N_age_die, aes(x=AGE, y=N_deaths)) +
  geom_line()
```

probability of death per year
```{r}
death_rate <- N_age2 %>%
  filter(AGE>=65) %>%
  left_join(N_age_die, by='AGE') %>%
  mutate( N_deaths = ifelse(is.na(N_deaths),0,N_deaths),
    death_rate= N_deaths/N_obs 
    ) 

  ggplot( death_rate, aes(x=AGE, y=death_rate)) +
  geom_line() +
    theme_classic()

```





 

