---
title: "1. exploratory analyses"
author: "Dan Weinberger"
date: '2023-01-05'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(arrow)
library(tidyr)
library(dplyr)
library(ggplot2)
library(duckdb)
library(plotly)
library(lubridate)
source('./R/schema.R')
state_cw <- read.csv('./Data/SSA_state_crosswalk.csv')
state.info <- cbind.data.frame(state.name, state.abb, state.region)

```

## Read in the parquet database

```{r} 
pq_stream_denom <- open_dataset("T:/parquet/denom/", format = "parquet", schema = sch_denom)
```

```{r} 
pq_stream_inp_k <- open_dataset("T:/parquet/inp_k/", format = "parquet" ) 
pq_stream_inp_j <- open_dataset("T:/parquet/inp_j/", format = "parquet")


```

```{r} 
pq_stream_snf <- open_dataset("T:/parquet/snf/", format = "parquet", schema = sch_inp_basek)
```

## Test databases 

Take a test pull from the dataset
```{r}
test1 <- pq_stream_denom %>% 
  slice_head( n = 10000) %>%
  collect() #use collect() to take parquet to dataframe


test2 <- pq_stream_denom %>% 
  filter(DSYSRTKY=='100009312') %>%
  collect() #use collect() to take parquet to dataframe


##DENOM ONLY HAS 2014-2016 DATA--AFTER THAT SWITCHES TO THE MBSF FILE
test3 <- pq_stream_denom %>%
  to_duckdb() %>%
  summarize(minyr=min(RFRNC_YR), maxyr=max(RFRNC_YR)) %>%
  collect()

```

Look at age distribution of all enrollees in 2015
should be ~54 million enrollees in 2015 (58.6 million in 2021)..

to run on the csv_stream takes 151 seconds
to run on the parquet data stream takes 22 seconds
```{r}
ptm <- proc.time()
N_age2 <- pq_stream_denom %>% 
  filter(AGE>=65 & RFRNC_YR==15) %>%
  select(AGE,RFRNC_YR) %>%
group_by(AGE) %>%
  summarize(N_obs = n()) %>% #use collect() to take parquet to dataframe
    collect()

proc.time() - ptm

ggplot(N_age2, aes(x=AGE, y=N_obs)) +
  geom_line()+
  theme_classic()
```


```{r}
test <-pq_stream_inp_k %>%
slice_head( n = 50) %>%
  collect() 
```

## Number of inpatient admissions by week, nationally

Break out by week and ICD chapter AND region
```{r,eval=F}
#runs in 2.5 min
ptm <- proc.time()
N_chapter_date_state <- pq_stream_inp_k %>%
  mutate(princ_icd_chapt =substr(ICD_DGNS_CD1,1,1)) %>%
  group_by(princ_icd_chapt,ADMSN_DT,STATE_CD) %>%
  summarize(N_cases = n()) %>%
  collect() %>%
  mutate(date=as.Date(ADMSN_DT,'%Y%m%d')) 
saveRDS(N_chapter_date_state,"T:/results/inp/N_chapter_date_state.rds")
proc.time() - ptm
```


```{r}
N_chapter_date_state <- readRDS("T:/results/inp/N_chapter_date_state.rds")

statedate <- N_chapter_date_state %>%
  filter(date>='2016-01-01' &date<='2021-11-20' )%>%
  mutate(STATE_CD=as.numeric(STATE_CD)) %>%
  left_join(state_cw, by=c('STATE_CD'='STATE_SSA')) %>%
left_join(state.info,by=c('STATENAME'='state.name')) %>%
  mutate(weekdate=lubridate::floor_date(date, 'week')) %>%
  group_by(princ_icd_chapt,state.region,weekdate) %>%
  summarize(N_cases=sum(N_cases)) %>%
  ungroup()
```

National, all causes
```{r}
statedate%>%
   group_by(weekdate) %>%
  summarize(N_cases=sum(N_cases)) %>%
    ggplot(aes(x=weekdate, y=N_cases))+
  geom_line()+
  theme_classic()+
  ylim(0,NA) +
  geom_vline(xintercept=as.Date(c('2020-03-15','2020-12-15','2021-08-15')), lty=2, col='gray')
```
mean Length of stay (LOS)
```{r}
ptm <- proc.time()
LOS_trends <- pq_stream_inp_k %>%
  mutate(LOS = UTIL_DAY + COIN_DAY) %>%
  group_by(ADMSN_DT) %>%
  summarize(LOS = mean(LOS)) %>%
  collect() %>%
  mutate(date=as.Date(ADMSN_DT,'%Y%m%d'), weekdate=lubridate::floor_date(date, 'week')) %>%  
  ungroup() %>%
  group_by(weekdate) %>%
  summarize(LOS = mean(LOS)) 
saveRDS(LOS_trends,"T:/results/inp/LOS_date.rds")

proc.time() - ptm


LOS_trends %>%
  filter(weekdate>='2017-01-01' &weekdate<='2021-11-20' )%>%
 ggplot(aes(x=weekdate, y=LOS))+
  geom_line()+
  theme_classic()+
  ylim(0,NA) +
  geom_vline(xintercept=as.Date(c('2020-03-15','2020-12-15','2021-08-15')), lty=2, col='gray')
```

National, by chapter
```{r, fig.width=10, fig.height=9}
statedate%>%
   group_by(princ_icd_chapt,weekdate) %>%
  summarize(N_cases=sum(N_cases)) %>%
    ggplot(aes(x=weekdate, y=N_cases))+
  facet_wrap(~princ_icd_chapt, scales='free_y') +
  geom_line()+
  theme_classic()+
  ylim(0,NA) +
  geom_vline(xintercept=as.Date(c('2020-03-15','2020-12-15','2021-08-15')), lty=2, col='gray')
```

Same but filter by region (Northeast, North Central, South, West
)
```{r}
statedate%>%
  filter(state.region=='West') %>%
  ggplot(aes(x=weekdate, y=N_cases))+
  geom_line()+
  facet_wrap(~princ_icd_chapt, scales='free_y') +
  theme_classic()+
  ylim(0,NA) +
  geom_vline(xintercept=as.Date(c('2020-03-15','2020-12-15','2021-08-15')), lty=2, col='gray')

```


NOTE This is currently ALL pneumonia; not just community-acquired (present on admission)

```{r}
ptm <- proc.time()
inp1 <- pq_stream_inp_k %>%
  mutate(date=as.Date(ADMSN_DT,'%Y%m%d'),     
             all_dx = paste0(ICD_DGNS_CD1,
                         ICD_DGNS_CD1,
                         ICD_DGNS_CD2,
                         ICD_DGNS_CD3,
                         ICD_DGNS_CD4,
                         ICD_DGNS_CD5,
                         ICD_DGNS_CD6,
                         ICD_DGNS_CD7,
                         ICD_DGNS_CD8,
                         ICD_DGNS_CD9,
                         ICD_DGNS_CD10,
                         ICD_DGNS_CD11,
                         ICD_DGNS_CD12,
                         ICD_DGNS_CD13,
                         ICD_DGNS_CD14,
                         ICD_DGNS_CD15,
                         ICD_DGNS_CD16,
                         ICD_DGNS_CD17,
                    ICD_DGNS_CD18,
                    ICD_DGNS_CD19,
                    ICD_DGNS_CD20,
                    ICD_DGNS_CD21,
                    ICD_DGNS_CD22,
                    ICD_DGNS_CD23,
                    ICD_DGNS_CD24,
                    ICD_DGNS_CD24)
                  )%>%
  filter( grepl('J12',all_dx)|grepl('J13',all_dx)|grepl('J14',all_dx)|grepl('J15',all_dx)|grepl('J16',all_dx)|grepl('J17',all_dx)|grepl('J18',all_dx)) %>%
  group_by(date) %>%
  summarize(J12_18=n()) %>%
  collect()
proc.time() - ptm

inp1 %>%
  arrange(date) %>%
  filter(date>='2017-01-01') %>%
  ggplot(aes( x=date, y=J12_18) )+
  geom_line()+
  theme_classic()
  
ptm <- proc.time()
#takes about 10 min to run
j12_18_line <- pq_stream_inp_k %>%
  #slice_head( n = 50) %>%
  mutate(#adate=as.Date(ADMSN_DT,'%Y%m%d'),
        # ddate = as.Date(DSCHRGDT,'%Y%m%d'),
             all_dx = paste(ICD_DGNS_CD1,
                         ICD_DGNS_CD1,
                         ICD_DGNS_CD2,
                         ICD_DGNS_CD3,
                         ICD_DGNS_CD4,
                         ICD_DGNS_CD5,
                         ICD_DGNS_CD6,
                         ICD_DGNS_CD7,
                         ICD_DGNS_CD8,
                         ICD_DGNS_CD9,
                         ICD_DGNS_CD10,
                         ICD_DGNS_CD11,
                         ICD_DGNS_CD12,
                         ICD_DGNS_CD13,
                         ICD_DGNS_CD14,
                         ICD_DGNS_CD15,
                         ICD_DGNS_CD16,
                         ICD_DGNS_CD17,
                    ICD_DGNS_CD18,
                    ICD_DGNS_CD19,
                    ICD_DGNS_CD20,
                    ICD_DGNS_CD21,
                    ICD_DGNS_CD22,
                    ICD_DGNS_CD23,
                    ICD_DGNS_CD24,
                    ICD_DGNS_CD24)
                  )%>%
  filter( grepl('J12',all_dx)|grepl('J13',all_dx)|grepl('J14',all_dx)|grepl('J15',all_dx)|grepl('J16',all_dx)|grepl('J17',all_dx)|grepl('J18',all_dx)) %>%
  select(DSYSRTKY, all_dx,ADMSN_DT,DSCHRGDT,GNDR_CD,RACE_CD, CNTY_CD, STATE_CD ,starts_with('CLM_POA_IND'),PTNTSTUS ) %>%
  collect()
proc.time() - ptm

j13_line <- j12_18_line %>%
  filter( grepl('J13',all_dx)) %>%
  mutate(adate=as.Date(ADMSN_DT,'%Y%m%d'),
         ddate=as.Date(DSCHRGDT,'%Y%m%d'),
         aweekdate=lubridate::floor_date(adate, 'week'),
         dweekdate=lubridate::floor_date(ddate, 'week'),
         LOS= as.numeric(difftime(ddate, adate, units='days' )),
         LOS=as.numeric(LOS),
         LOS_grp = if_else(LOS %in% c(0,1),1,
                        if_else(LOS>=2 & LOS<=6,2,
                        if_else(LOS>=7 & LOS<=13,3,
                        if_else(LOS>=14 ,4,999))))
  )

hist(j13_line$LOS)

#Group by alive/dead
j13_line%>%
  group_by(PTNTSTUS,aweekdate) %>%
  summarize(N_cases=n()) %>%
  filter(PTNTSTUS %in% c('A','B')) %>%
  ggplot(aes(x=aweekdate, y=N_cases))+
  geom_line() +
  theme_classic()+
  facet_wrap(~PTNTSTUS, scales='free_y' )+
  ggtitle('Trend in those that die vs survive J13')

#Group by LOS

j13_line%>%
  group_by(LOS_grp,aweekdate) %>%
  summarize(N_cases=n()) %>%
  filter(LOS_grp %in% c(1,2,3,4)) %>%
  mutate(LOS_grp=as.factor(LOS_grp)) %>%
  ggplot(aes(x=aweekdate, y=N_cases))+
  geom_line() +
  theme_classic()+
  facet_wrap(~LOS_grp, scales='free_y' )+
  ylim(0,NA) +
  ggtitle('Trend in long vs short duration stay, J13')

j13_line%>%
  filter(PTNTSTUS %in% c('A','B') &
           !grepl('A40',all_dx) & !grepl('A41',all_dx)) %>%
  group_by(PTNTSTUS,aweekdate) %>%
  summarize(N_cases=n()) %>%
  
  ggplot(aes(x=aweekdate, y=N_cases))+
  geom_line() +
  theme_classic()+
  facet_wrap(~PTNTSTUS, scales='free_y' )

```

pneumococcal codes 
```{r}
ptm <- proc.time()
pneumo.line.list <- pq_stream_inp_k %>%
  mutate(date=as.Date(ADMSN_DT,'%Y%m%d'),     
             all_dx = paste0(ICD_DGNS_CD1,
                         ICD_DGNS_CD1,
                         ICD_DGNS_CD2,
                         ICD_DGNS_CD3,
                         ICD_DGNS_CD4,
                         ICD_DGNS_CD5,
                         ICD_DGNS_CD6,
                         ICD_DGNS_CD7,
                         ICD_DGNS_CD8,
                         ICD_DGNS_CD9,
                         ICD_DGNS_CD10,
                         ICD_DGNS_CD11,
                         ICD_DGNS_CD12,
                         ICD_DGNS_CD13,
                         ICD_DGNS_CD14,
                         ICD_DGNS_CD15,
                         ICD_DGNS_CD16,
                         ICD_DGNS_CD17,
                    ICD_DGNS_CD18,
                    ICD_DGNS_CD19,
                    ICD_DGNS_CD20,
                    ICD_DGNS_CD21,
                    ICD_DGNS_CD22,
                    ICD_DGNS_CD23,
                    ICD_DGNS_CD24,
                    ICD_DGNS_CD24)
                  )%>%
    select(DSYSRTKY, all_dx,ADMSN_DT,DSCHRGDT,RACE_CD,  STATE_CD  ) %>%
  filter( grepl('J13',all_dx)|grepl('A403',all_dx)|grepl('G001',all_dx)) %>%
  collect()
proc.time() - ptm


p1 <- pneumo.line.list %>%
 mutate( adate=as.Date(ADMSN_DT,'%Y%m%d'),
         aweekdate=lubridate::floor_date(adate, 'week')
) %>%  
  filter(aweekdate >= '2016-01-01' & aweekdate<='2021-11-30') %>%
  group_by(aweekdate) %>%
  summarize(N=n()) %>%
  ggplot(aes(x=aweekdate, y=N)) +
  theme_classic()+
  geom_line()+
  ylim(0,NA)

p1

## multiple records per person
pneumo.line.list %>%
  group_by(DSYSRTKY) %>%
  summarize(N_records=n()) %>%
  ungroup() %>%
  group_by(N_records) %>%
  summarize(N=n())


```


RSV line list
```{r}
rsv_line_list_k <- pq_stream_inp_k %>%
  mutate(adate=as.Date(ADMSN_DT,'%Y%m%d'),     
             all_dx = paste(ICD_DGNS_CD1,
                         ICD_DGNS_CD1,
                         ICD_DGNS_CD2,
                         ICD_DGNS_CD3,
                         ICD_DGNS_CD4,
                         ICD_DGNS_CD5,
                         ICD_DGNS_CD6,
                         ICD_DGNS_CD7,
                         ICD_DGNS_CD8,
                         ICD_DGNS_CD9,
                         ICD_DGNS_CD10,
                         ICD_DGNS_CD11,
                         ICD_DGNS_CD12,
                         ICD_DGNS_CD13,
                         ICD_DGNS_CD14,
                         ICD_DGNS_CD15,
                         ICD_DGNS_CD16,
                         ICD_DGNS_CD17,
                    ICD_DGNS_CD18,
                    ICD_DGNS_CD19,
                    ICD_DGNS_CD20,
                    ICD_DGNS_CD21,
                    ICD_DGNS_CD22,
                    ICD_DGNS_CD23,
                    ICD_DGNS_CD24,
                    ICD_DGNS_CD24)
                  )%>%
 filter(grepl("B974|J121|J205|J210", all_dx)) %>%
  select(DSYSRTKY,all_dx,adate,STATE_CD,CNTY_CD,TOT_CHRG) %>%
  collect()

rsv_line_list_j <- pq_stream_inp_j %>%
   rename_with(toupper) %>%
  mutate(adate=as.Date(CLM_ADMSN_DT,'%Y%m%d'),     
             all_dx = paste(ICD_DGNS_CD1,
                         ICD_DGNS_CD1,
                         ICD_DGNS_CD2,
                         ICD_DGNS_CD3,
                         ICD_DGNS_CD4,
                         ICD_DGNS_CD5,
                         ICD_DGNS_CD6,
                         ICD_DGNS_CD7,
                         ICD_DGNS_CD8,
                         ICD_DGNS_CD9,
                         ICD_DGNS_CD10,
                         ICD_DGNS_CD11,
                         ICD_DGNS_CD12,
                         ICD_DGNS_CD13,
                         ICD_DGNS_CD14,
                         ICD_DGNS_CD15,
                         ICD_DGNS_CD16,
                         ICD_DGNS_CD17,
                    ICD_DGNS_CD18,
                    ICD_DGNS_CD19,
                    ICD_DGNS_CD20,
                    ICD_DGNS_CD21,
                    ICD_DGNS_CD22,
                    ICD_DGNS_CD23,
                    ICD_DGNS_CD24,
                    ICD_DGNS_CD24,sep='_')
                  )%>%
 filter(grepl("B974|J121|J205|J210|0796|46611|4801", all_dx)) %>%
  select(DESY_SORT_KEY,all_dx,adate,BENE_STATE_CD,BENE_CNTY_CD,CLM_TOT_CHRG_AMT) %>%
  collect() %>%
  rename(DSYSRTKY=DESY_SORT_KEY,
         STATE_CD=BENE_STATE_CD,
         CNTY_CD = BENE_CNTY_CD,
         TOT_CHRG = CLM_TOT_CHRG_AMT) %>%
  mutate(DSYSRTKY=as.numeric(DSYSRTKY))

rsv_line_list <- bind_rows(rsv_line_list_j, rsv_line_list_k)

saveRDS(rsv_line_list,'T:/results/inp/rsv_id_dates.rds')

```

```{r, fig.width=3.5, fig.height=3}
rsv_line_list <- readRDS('T:/results/inp/rsv_id_dates.rds')


#library(cdcfluview)
#data(hhs_regions)

rsv1 <- rsv_line_list %>%
  mutate(weekdate=floor_date(adate, 'week')) %>%
 # filter(weekdate>='2017-01-01') %>%
  group_by(STATE_CD,weekdate) %>%
  summarize(N_rsv = n() ) %>%
  mutate(STATE_CD=as.numeric(STATE_CD),
         week=week(weekdate),
         year=year(weekdate),
         epiyr = if_else(week>=27,year, year-1)) %>%
  ungroup() %>%
  group_by(STATE_CD, epiyr) %>%
  mutate( RSV_adj = N_rsv/sum(N_rsv)) %>%
  ungroup() %>%
  group_by(STATE_CD) %>%
  mutate(       mov_ave_RSV = zoo::rollapply(N_rsv, align='center',FUN=mean, width=53,  partial=T),
                RSV_adj_movave =N_rsv/mov_ave_RSV ) %>%
  left_join(state_cw, by=c('STATE_CD'='STATE_SSA'))  %>%
  mutate(STATENAME = if_else( STATENAME=='New York', 'NY', STATENAME))

rsv1 %>%
  filter(STATENAME %in% c('Florida','NY') & epiyr>=2014) %>%
ggplot( aes(x=weekdate, y=RSV_adj_movave, group=STATENAME, color=STATENAME)) +
  geom_line(show.legend = FALSE)+
  theme_classic()+
  ylab('RSV Hospitalizations (adjusted)')+
  xlab('') +
    directlabels::geom_dl(aes(label = STATENAME), method = "smart.grid")

rsv1 %>%
  filter(STATENAME %in% c('Florida','NY') & epiyr>=2014) %>%
ggplot( aes(x=weekdate, y=N_rsv, group=STATENAME, color=STATENAME)) +
  geom_line(show.legend = FALSE)+
  theme_classic()+
  ylab('RSV Hospitalizations (unadjusted)')+
  xlab('') +
    directlabels::geom_dl(aes(label = STATENAME), method = "smart.grid")


ggplot(rsv_line_list , aes(x=log(TOT_CHRG))) +
  geom_histogram()+
  theme_classic()+
  ylab('charges')


rsv1 <- rsv_line_list %>%
  mutate(weekdate=floor_date(adate, 'week')) %>%
  
 # filter(weekdate>='2017-01-01') %>%
  group_by(weekdate) %>%
  summarize(N_rsv = n() )
```


```{r}
pneumonia_line_list <- pq_stream_inp_k %>%
  mutate(adate=as.Date(ADMSN_DT,'%Y%m%d'),     
             all_dx = paste0(ICD_DGNS_CD1,
                         ICD_DGNS_CD1,
                         ICD_DGNS_CD2,
                         ICD_DGNS_CD3,
                         ICD_DGNS_CD4,
                         ICD_DGNS_CD5,
                         ICD_DGNS_CD6,
                         ICD_DGNS_CD7,
                         ICD_DGNS_CD8,
                         ICD_DGNS_CD9,
                         ICD_DGNS_CD10,
                         ICD_DGNS_CD11,
                         ICD_DGNS_CD12,
                         ICD_DGNS_CD13,
                         ICD_DGNS_CD14,
                         ICD_DGNS_CD15,
                         ICD_DGNS_CD16,
                         ICD_DGNS_CD17,
                    ICD_DGNS_CD18,
                    ICD_DGNS_CD19,
                    ICD_DGNS_CD20,
                    ICD_DGNS_CD21,
                    ICD_DGNS_CD22,
                    ICD_DGNS_CD23,
                    ICD_DGNS_CD24,
                    ICD_DGNS_CD24)
                  )%>%
  filter( grepl('J12',all_dx)|grepl('J13',all_dx)|grepl('J14',all_dx)|grepl('J15',all_dx)|grepl('J16',all_dx)|grepl('J17',all_dx)|grepl('J18',all_dx)) %>%
  select(DSYSRTKY,all_dx,adate) %>%
  collect()

saveRDS(pneumonia_line_list,'T:/results/inp/J12_j18_id_dates.rds')

any_return_hosp <- pq_stream_inp_k %>%
  mutate(adate2=as.Date(ADMSN_DT,'%Y%m%d')) %>%     
  right_join(pneumonia_line_list , by='DSYSRTKY') %>%
    filter(adate2 > adate) %>% #hospitalizations that occur AFTER pneumonia hospitalization
    summarize(N_events=n()) %>%
    collect()
  

pneumonia_line_list <- readRDS('T:/results/inp/J12_j18_id_dates.rds')

pneu1 <- pneumonia_line_list %>%
  mutate(weekdate=floor_date(adate, 'week')) %>%
  group_by(weekdate) %>%
  summarize(N_pneu = n() )

ggplot(pneu1 , aes(x=weekdate, y=N_pneu)) +
  geom_line()+
  theme_classic()
  
```



```{r}
ptm <- proc.time()
N_age_die <- pq_stream_denom %>%
  filter(DEATH_DT != '') %>%
  select(AGE,RFRNC_YR) %>%
  filter(RFRNC_YR==15) %>%
group_by(AGE) %>%
  summarize(N_deaths = n()) %>% #use collect() to take parquet to dataframe
    collect()

proc.time() - ptm

ggplot(N_age_die, aes(x=AGE, y=N_deaths)) +
  geom_line()
```

probability of death per year
```{r}
death_rate <- N_age2 %>%
  filter(AGE>=65) %>%
  left_join(N_age_die, by='AGE') %>%
  mutate( N_deaths = ifelse(is.na(N_deaths),0,N_deaths),
    death_rate= N_deaths/N_obs 
    ) 

  ggplot( death_rate, aes(x=AGE, y=death_rate)) +
  geom_line() +
    theme_classic()

```


## skilled nursing trend

test pull
```{r}
test2 <- pq_stream_snf %>% 
  slice_head( n = 100) %>%
  collect() #use collect() to take parquet to dataframe
```

```{r,eval=F}
#20151210 ADMSN_DT  WATCH OUT FOR DUPLICATES  ADMSN_DT DSCHRGDT
#runs ~5 min
ptm <- proc.time()
N_snf_date_state <- pq_stream_snf |> 
    to_duckdb() |> #convert to duckdb database to allow for row_number
  select(DSYSRTKY,ADMSN_DT,ADMSN_DT,STATE_CD,DSCHRGDT) %>%
  group_by(DSYSRTKY,ADMSN_DT) %>%
  collect() %>%
  mutate(record_N=row_number()) %>% #removes duplicates
  ungroup() %>%
  filter(record_N==1) %>%
  group_by(ADMSN_DT,STATE_CD) %>%
  summarize(N_cases = n()) %>%
  #collect() %>%
  mutate(date=as.Date(ADMSN_DT,'%Y%m%d')) 

saveRDS(N_snf_date_state,"T:/results/snf/N_date_state_snf.rds")
proc.time() - ptm
```

```{r}
statedate_snf <- N_snf_date_state %>%
  filter(date>='2017-01-01' &date<='2021-11-20' )%>%
  mutate(STATE_CD=as.numeric(STATE_CD)) %>%
  left_join(state_cw, by=c('STATE_CD'='STATE_SSA')) %>%
left_join(state.info,by=c('STATENAME'='state.name')) %>%
  mutate(weekdate=lubridate::floor_date(date, 'week')) %>%
  group_by(state.region,weekdate) %>%
  summarize(N_cases=sum(N_cases)) %>%
  ungroup()

```


```{r}
statedate_snf%>%
   group_by(weekdate) %>%
  summarize(N_cases=sum(N_cases)) %>%
    ggplot(aes(x=weekdate, y=N_cases))+
  geom_line()+
  theme_classic()+
  ylim(0,NA) +
  geom_vline(xintercept=as.Date(c('2020-03-15','2020-12-15','2021-08-15')), lty=2, col='gray')
```

LOS in snf

45 sec!
```{r}
ptm <- proc.time() 
LOS_date_admit <- pq_stream_snf %>% 
  to_duckdb() %>%
  filter(DSCHRGDT !='') %>%
   mutate(
         dateADMT=strptime(ADMSN_DT,'%Y%m%d'),
         dateDIS=strptime(DSCHRGDT,'%Y%m%d'),
         LOS2 = datediff('day',dateADMT,dateDIS)
         ) %>%
  group_by(dateADMT) %>%
  filter(dateADMT>='2016-01-01' & dateADMT<='2021-11-20' ) %>%
  select(dateADMT, LOS2) %>%
  collect() 

proc.time() - ptm


```

```{r}
LOS_date_admit%>%
   mutate(aweekdate= lubridate::floor_date(dateADMT,'week')) %>%
  ungroup() %>%
      group_by(aweekdate) %>% 
    summarize(LOS2=mean(LOS2)) %>%
  ggplot(aes(x=aweekdate, y=LOS2))+
  geom_line()+
  theme_classic()
```
### Occupancy
SNF
```{r}
los_date_admit_summary <- LOS_date_admit%>%
    group_by(dateADMT, LOS2) %>%
   summarize(N=n()) %>%
  mutate(dateADMT=as.Date(dateADMT),
         dateDIS= dateADMT + LOS2)

all.date <- seq.Date(from=min(los_date_admit_summary$dateADMT), to=max(los_date_admit_summary$dateADMT), by='week')


occupancy <- pbapply::pblapply(all.date, function(x){
  occ <-   los_date_admit_summary %>%
    ungroup() %>%
    filter(dateADMT<=x & dateDIS>=x) %>%
    summarize(occupancy=sum(N))  
  return(occ)
}) %>% 
  bind_rows() 

cbind.data.frame(all.date,occupancy) %>%
  filter(all.date>='2016-03-01') %>%
ggplot( aes(x=all.date , y=occupancy)) +
  geom_line() +
  ylim(0, NA) +
  theme_classic()

```
Inpatients
```{r}
ptm <- proc.time() 
LOS_date_admit_inp <- pq_stream_inp_k %>% 
  to_duckdb() %>%
  filter(DSCHRGDT !='') %>%
   mutate(
         dateADMT=strptime(ADMSN_DT,'%Y%m%d'),
         dateDIS=strptime(DSCHRGDT,'%Y%m%d'),
         LOS2 = datediff('day',dateADMT,dateDIS)
         ) %>%
  group_by(dateADMT) %>%
  filter(dateADMT>='2016-01-01' & dateADMT<='2021-11-20' ) %>%
  select(dateADMT, LOS2) %>%
  collect() 

LOS_date_admit_inp_summary <- LOS_date_admit_inp %>%
group_by(dateADMT, LOS2) %>%
   summarize(N=n()) %>%
  mutate(dateADMT=as.Date(dateADMT),
         dateDIS= dateADMT + LOS2)

all.date_inp <- seq.Date(from=min(LOS_date_admit_inp_summary$dateADMT), to=max(LOS_date_admit_inp_summary$dateADMT), by='week')


occupancy_inp <- pbapply::pblapply(all.date, function(x){
  occ <-   LOS_date_admit_inp_summary %>%
    ungroup() %>%
    filter(dateADMT<=x & dateDIS>=x) %>%
    summarize(occupancy=sum(N))  
  return(occ)
}) %>% 
  bind_rows() 

cbind.data.frame(all.date_inp,occupancy_inp) %>%
  filter(all.date>='2016-03-01') %>%
ggplot( aes(x=all.date_inp , y=occupancy)) +
  geom_line() +
  ylim(0, NA) +
  theme_classic()
proc.time() - ptm


```
RESTRICT TO AZ (state=3)
```{r}

ptm <- proc.time() 
LOS_date_admit_inp_AZ <- pq_stream_inp_k %>% 
  to_duckdb() %>%
  filter(DSCHRGDT !='' & STATE_CD == '03' ) %>%
   mutate(
         dateADMT=strptime(ADMSN_DT,'%Y%m%d'),
         dateDIS=strptime(DSCHRGDT,'%Y%m%d'),
         LOS2 = datediff('day',dateADMT,dateDIS)
         ) %>%
  group_by(dateADMT) %>%
  filter(dateADMT>='2016-01-01' & dateADMT<='2021-11-20' ) %>%
  select(dateADMT, LOS2) %>%
  collect() 

LOS_date_admit_inp_summary_AZ <- LOS_date_admit_inp_AZ %>%
group_by(dateADMT, LOS2) %>%
   summarize(N=n()) %>%
  mutate(dateADMT=as.Date(dateADMT),
         dateDIS= dateADMT + LOS2)

all.date_inp_AZ <- seq.Date(from=min(LOS_date_admit_inp_summary_AZ$dateADMT), to=max(LOS_date_admit_inp_summary_AZ$dateADMT), by='week')


occupancy_inp_AZ <- pbapply::pblapply(all.date_inp_AZ, function(x){
  occ <-   LOS_date_admit_inp_summary_AZ %>%
    ungroup() %>%
    filter(dateADMT<=x & dateDIS>=x) %>%
    summarize(occupancy=sum(N))  
  return(occ)
}) %>% 
  bind_rows() 

p1 <-cbind.data.frame(all.date_inp_AZ,occupancy_inp_AZ) %>%
  filter(all.date_inp_AZ>='2017-03-01') %>%
ggplot( aes(x=all.date_inp_AZ , y=occupancy)) +
  geom_line() +
  ylim(0, NA) +
  theme_classic()
p1
ggplotly(p1)
proc.time() - ptm
```

```{r}

```


 

